Brownian motion on an upswitch

### 2017-06-14 examples/oscillator.Rmd
upswitch example

```{r setup, include=FALSE}
library(expm)
library(knitr)
library(parallel)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
```

@@

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix}, \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} 
\end{aligned}$$
and
$$\begin{aligned}
    A = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix} .
\end{aligned}$$
This system has transfer function
$$\begin{aligned}
    h(t) &= C e^{tA} B \\
        &= sin(t) + cos(t) .
\end{aligned}$$
The set of equivalent $A$ are given by
$\{A(\tau): \tau \neq 1 \in \R\}$, where
$$\begin{aligned}
    V(\tau) 
    &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A(\tau) 
    &= V(\tau) A V(\tau)^{-1} \\
    % &= \begin{bmatrix} 0 & 1 \\ \tau-1 & \tau \end{bmatrix} 
    %    \begin{bmatrix} 1 & 0 \\ -\tau/(1-\tau) & 1/(1-\tau) \end{bmatrix}  \\
    &= \begin{bmatrix} -\tau/(1-\tau) & 1/(1-\tau) \\ -(\tau^2 + (1-\tau)^2)/(1-\tau) & \tau/(1-\tau) \end{bmatrix} 
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

```{r system_setup}
A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)
V <- function (tau) { matrix(c(1,tau,0,1-tau), nrow=2) }
Vinv <- function (tau) { matrix(c(1,-tau/(1-tau),0,1/(1-tau)), nrow=2) }
At <- function (tau) { V(tau) %*% A0 %*% Vinv(tau) }
A_fn <- function (tau) { 
    matrix(c(-tau/(1-tau), -(tau^2+(1-tau)^2)/(1-tau), 
             1/(1-tau), tau/(1-tau)), nrow=2) 
}
h <- function (t,tau=0,A=A_fn(tau)) { sapply(t, function (tt) C %*% expm(tt*A) %*% B) }
```

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, sapply(tt, h), type='l',
     xlab='t', ylab=expression(h(t)), main='impulse response function')
lines(tt, sin(tt)+cos(tt), col='red', lty=2)
```

@@

```{r distance_fun}
D <- function (A, upper=100, ...) {
    f <- function (t) {
        exp(-t/(4*pi)) * ( h(t,A=A) - cos(t) - sin(t) )^2
    }
    integrate(f, lower=0, upper=upper, ...)$value
}
```

To check this works, the distance of $A(2)$ to optimal should be zero:
```{r check_D}
D(A(2))
```

@@

Here are the phenotypes for $\tau=2$, $\tau=2.01$, and the hybrids:
```{r phenotypes}
dtau <- .01
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")
```
And, the same thing for $\tau=2.1$:
```{r phenotypes2}
dtau <- .1
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")
```

@@

>>>>

```{r setup, include=FALSE}
library(expm)
library(knitr)
library(parallel)
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
do.pdfs <- TRUE
pdf_copy <- function ( width=6, height=width*knitr::opts_current$get("fig.height")/knitr::opts_current$get("fig.width"), plot.id=NULL,
                filename) {
    if (missing(filename)) {
        file.id <- if (is.null(plot.id)) { "" } else { paste0("_",plot.id) }
        filename <- knitr::fig_path(paste(file.id,".pdf",sep=""))
    }
    cat("pdf version at:",filename)
    dev.print( file=filename, device=pdf, width=width, height=height, pointsize=10, family="sans")
}
```

@@

Consider the dynamical system given by
$$\begin{aligned}
      \dot{x}(t) &= Ax(t) + Bu(t) \\
          y(t) &= Cx(t) 
\end{aligned}$$
where
$$\begin{aligned}
    B^{T} = \begin{bmatrix} 1 & 1 \end{bmatrix}, \qquad C = \begin{bmatrix} 1 & 0 \end{bmatrix} 
\end{aligned}$$
and
$$\begin{aligned}
    A = \begin{bmatrix} 0 & 1 \\ -1 & 0 \end{bmatrix} .
\end{aligned}$$
This system has impulse response function
$$\begin{aligned}
    h(t) &= C e^{tA} B \\
        &= sin(t) + cos(t) .
\end{aligned}$$
The set of equivalent $A$ are given by
$\{A(\tau): \tau \neq 1 \in \R\}$, where
$$\begin{aligned}
    V(\tau) 
    &= \begin{bmatrix} 1 & 0 \\ \tau & 1-\tau \end{bmatrix} \\
    A(\tau) 
    &= V(\tau) A V(\tau)^{-1} \\
    % &= \begin{bmatrix} 0 & 1 \\ \tau-1 & \tau \end{bmatrix} 
    %    \begin{bmatrix} 1 & 0 \\ -\tau/(1-\tau) & 1/(1-\tau) \end{bmatrix}  \\
    &= \begin{bmatrix} -\tau/(1-\tau) & 1/(1-\tau) \\ -(\tau^2 + (1-\tau)^2)/(1-\tau) & \tau/(1-\tau) \end{bmatrix} 
\end{aligned}$$
These systems are input-output equivalent (although may have different noise properties).

```{r system_setup}
A0 <- matrix(c(0,-1,1,0), nrow=2)
B <- matrix(c(1,1), ncol=1)
C <- matrix(c(1,0), nrow=1)
V <- function (tau) { matrix(c(1,tau,0,1-tau), nrow=2) }
Vinv <- function (tau) { matrix(c(1,-tau/(1-tau),0,1/(1-tau)), nrow=2) }
At <- function (tau) { V(tau) %*% A0 %*% Vinv(tau) }
A_fn <- function (tau) { 
    matrix(c(-tau/(1-tau), -(tau^2+(1-tau)^2)/(1-tau), 
             1/(1-tau), tau/(1-tau)), nrow=2) 
}
h <- function (t,tau=0,A=A_fn(tau)) { sapply(t, function (tt) C %*% expm(tt*A) %*% B) }
optimal_h <- function (t) { sin(t)+cos(t) }
```

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, h(tt), type='l',
     xlab='t', ylab=expression(h(t)), main='impulse response function')
lines(tt, h(tt, tau=0.1), col='blue', lty=3)
lines(tt, optimal_h(tt), col='red', lty=2)

if (do.pdfs) { pdf_copy() }
```

@@

```{r distance_fun}
Df <- function (A,t) {
    exp(-t/(4*pi)) * ( h(t,A=A) - optimal_h(t) )^2
}
D <- function (A, upper=10, ...) {
    f <- function (t) { Df(A,t) }
    integrate(f, lower=0, upper=upper, ...)$value
}
```

To check this works, the distance of $A(2)$ to optimal should be zero:
```{r check_D}
D(A_fn(2))
```

@@

Here are the phenotypes for $\tau=2$, $\tau=2.01$, and the hybrids:
```{r phenotypes}
dtau <- .01
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")

if (do.pdfs) { pdf_copy() }
```
And, the same thing for $\tau=2.1$:
```{r phenotypes2}
dtau <- .1
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype")
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype")

if (do.pdfs) { pdf_copy() }
```

@@

if (do.pdfs) { pdf_copy() }
```

### 2017-06-14 examples/upswitch.Rmd
upswitch example

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, h(tt), type='l',
     xlab='t', ylab=expression(h(t)), main='impulse response function')
lines(tt, h(tt, tau=0.1), col='blue', lty=3)
lines(tt, optimal_h(tt), col='red', lty=2)
```

@@

Here are the phenotypes for $\tau=2$, $\tau=2.01$, and the hybrids:
```{r phenotypes}
dtau <- .01
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype", log='y')
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype", log='y')
```
And, the same thing for $\tau=2.1$:
```{r phenotypes2}
dtau <- .1
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype", log='y')
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype", log='y')
```

>>>>

Here is the impulse response function:
```{r impulse_response_plot}
tt <- seq(0,2*pi, length.out=100)
plot(tt, h(tt), type='l',
     xlab='t', ylab=expression(h(t)), main='impulse response function')
lines(tt, h(tt, tau=0.1), col='blue', lty=3)
lines(tt, optimal_h(tt), col='red', lty=2)

if (do.pdfs) { pdf_copy() }
```

@@

Here are the phenotypes for $\tau=2$, $\tau=2.01$, and the hybrids:
```{r phenotypes}
dtau <- .01
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype", log='y')
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype", log='y')

if (do.pdfs) { pdf_copy() }
```
And, the same thing for $\tau=2.1$:
```{r phenotypes2}
dtau <- .1
tt <- seq(0,6*2*pi, length.out=500)
hh <- sapply( list(A_fn(2), A_fn(2+dtau), F1(2, 2+dtau)), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main=sprintf("Parental phenotypes and F1s at dtau=%s", dtau), lty=c(1,1,2), col=1:3, xlab="time", ylab="phenotype", log='y')
legend("topright", lty=c(1,1,2), col=1:3,
        legend=c(expression(tau==2), expression(tau==2+dtau), "F1"))
# now the F2s
hh <- sapply( all_F2s(2, 2+dtau), function (A) { sapply(tt, h, A=A) } )
matplot(tt, hh, type='l', main="F2s", xlab="time", ylab="phenotype", log='y')

if (do.pdfs) { pdf_copy() }
```
